---
title: "Heart"
author: "Kinjal Mehta"
date: "June 7, 2018"
output:
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set seed}
# set seed
set.seed(123)
```

```{r set work directory and read the data}
setwd("C:/Users/kinja/Desktop/HS 628/Assignments/hs628_assign2_Data_Cleaning_and_Visualization_in_R/")
heart <- read.csv("heart.csv", encoding = "UTF-16", header = TRUE)
```

```{r}
class(heart)
str(heart)
```

The data type of heart is data.frame.

```{r}
table(heart$age)
table(heart$trestbps)
table(heart$chol)
```



```{r}
table(heart$oldpeak)
table(heart$thalach)
table(heart$ca)
table(heart$num)
```


```{r}
table(heart$cp)
table(heart$restecg)
table(heart$slope)
table(heart$thal)
```


```{r}
table(heart$sex)
table(heart$fbs)
table(heart$exang)
```


```{r}
heart
```



```{r correct variable type}
heart <- read.csv('heart.csv', encoding = "UTF-16", header = TRUE, na.string="?")
```



```{r summary of the data}
summary(heart)
```

```{r}
str(heart)
```

Write a couple of sentences on your observations.

```{r missing data}
heart <- read.csv('heart.csv', encoding = "UTF-16", header = TRUE, na.string=c("?", "-9"))
```

```{r}
summary(heart)
```

```{r}
str(heart)
```



```{r}
fit_num_all = lm(num ~ ., data = heart)
summary(fit_num_all)
```



```{r}
summary(na.omit(heart))
```

```{r}
nrow(na.omit(heart))
```

How many observations were deleted due to
missingness?

How many are left in the model?

Could this many points (observations) determine even a 2D (2-dimensional) line?

```{r}
colnames(heart)
```


```{r}
fit_num_1 = lm(num ~ age +sex +cp +trestbps +chol +fbs +restecg +thalach +exang +oldpeak +slope +thal, data = heart)
summary(fit_num_1)
```

How many observations are included in the
model? Could this many points determine a 13D hyperplane?

```{r}
fit_num_2 = lm(num ~ age +sex +cp +trestbps +chol +fbs +restecg +thalach +exang +oldpeak +slope, data = heart)
summary(fit_num_2)
```

How many observations are included in the
model? Could this many points determine a 12D hyperplane?

```{r Modify the non-boolean categorical predictor variables}
heart$cp <- as.factor(heart$cp)
heart$restecg <- as.factor(heart$restecg)
heart$slope <- as.factor(heart$slope)
heart$thal <- as.factor(heart$thal)
str(heart)
```

List the categorical variables that are
boolean. 

List the other categorical variables (those that can hold more than two values).
num, 

What type did R assign to the categorical variables?

```{r }
fit_num_3 = lm(num ~ age +sex +cp +trestbps +chol +fbs +restecg +thalach +exang +oldpeak +slope, data = heart)
summary(fit_num_3)
```

Execute a summary of fit_num_3 & compare it to your summary of fit_num_2, which has 11 predictor variables with estimates of slopes in the fitted model.

How has the list of fitted predictor variables changed in fit_num_3?

How has the fit of the significant variables changed?

```{r Modify the boolean categorical variables}
heart$sex <- as.factor(heart$sex)
heart$fbs <- as.factor(heart$fbs)
heart$exang <- as.factor(heart$exang)
str(heart)
```


```{r}
fit_num_4 = lm(num ~ age +sex +cp +trestbps +chol +fbs +restecg +thalach +exang +oldpeak +slope, data = heart)
summary(fit_num_4)
```

Execute a summary of fit_num_4 & compare it to your summary of fit_num_3.

How has the list of fitted predictor variables changed in fit_num_4?

```{r}
heart$num2[heart$num == 0] <- 0
heart$num2[heart$num != 0] <- 1
```



```{r}
fit_num_slope_oldpeak_sex <- lm(num ~ slope + oldpeak + sex, data = heart)
summary(fit_num_slope_oldpeak_sex)
```



```{r}
fit_num2_slope_oldpeak_sex <- lm(num2 ~ slope + oldpeak + sex, data = heart)
summary(fit_num2_slope_oldpeak_sex)
```

What is the equation of the fitted line for fit_num2_slope_oldpeak_sex?

How does the adjusted R2 of fit_num2_slope_oldpeak_sex compare with the
adjusted R2 of fit_num_slope_oldpeak_sex?

What does this indicate to you?

```{r}
plot(fit_num_1, 1)
plot(fit_num_slope_oldpeak_sex, 1)
plot(fit_num2_slope_oldpeak_sex, 1)
```

The plots produced by these commands indicate problems with these linear regression models. (See Assumption (iv) normality of the error distribution)

Write a couple of sentences identifying the problems.

